---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data
```{r}
library(dplyr)

activity <- read.csv("activity.csv")
activity_comp <- activity[complete.cases(activity),]

```

## What is mean total number of steps taken per day?
```{r}
activity_sum <- activity_comp %>%
  group_by(date) %>%
    summarise(count = sum(steps), mean_steps= mean(steps), median_steps=median(steps))
```

###Histogram 
```{r}

hist(activity_sum$count, xlab="No.Steps (Day)", ylab="Frequency", main="Histogram of Steps/Day", col="steelblue")

```

##Mean and median total number of steps taken per day

```{r}
summary <- activity_sum %>% 
                  select(date, mean_steps, median_steps)

summary
```

##Average daily activity pattern
```{r}
library(ggplot2)
activity_interval <- activity_comp %>%
  group_by(interval) %>%
    summarise(mean_steps= mean(steps))

max <- max(activity_interval$mean_steps) # Find a dplyr way to do this.
max_interval <- activity_interval %>% filter(mean_steps == max) %>% select(interval)

ggplot(activity_interval,aes(interval, mean_steps)) +
  geom_line() 
#+ geom_text(data=activity_interval[max_interval, max], label="note" , vjust=1, colour="red")


```

######5-minute interval that has the average maximum number of steps is `r max_interval`


##Imputing missing values
```{r}
activity2 <- read.csv("activity.csv")
total_nas <- sum(is.na(activity2$steps))

activity2[is.na(activity2)] <- mean(activity2$steps, na.rm = TRUE)
str(activity2)

library(ggplot2)

activity2_sum <- activity2 %>%
  group_by(date) %>%
    summarise(count = sum(steps), mean_steps= mean(steps), median_steps=median(steps))


hist(activity2_sum$count, xlab="No.Steps (Day)", ylab="Frequency", main="Histogram of Steps/Day", col="steelblue")

summary <- activity2_sum %>% select(date, mean_steps, median_steps)

summary

```
Total missing values in the dataset are `r total_nas`. The overall shape of the histogram remains the same with the exception of the Frequency being higher on the imputed values (imputed with the overall mean). 

### Comparing the differences between the original and imputed data sets
```{r}
imp <- select(activity2_sum, date, count )
colnames(imp) <- c("date", "imp_count")

orig <- select(activity_sum, date, count )
colnames(orig) <- c("date", "orig_count")

join <- left_join(imp, orig, by="date")

mutate <- mutate(join, diff=(imp_count-orig_count))

```
###The difference between original and imputed values
```{r}
mutate
```
###Weekdays vs Weekends
```{r}
activity2 <- read.csv("activity.csv")
activity2[is.na(activity2)] <- mean(activity2$steps, na.rm = TRUE)

activity2$date <- as.Date(activity2$date)

activity2 <- mutate(activity2, weekday = as.factor(ifelse(weekdays(activity2$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")))

activity2_weekend <- activity2 %>%
  group_by(weekday, interval) %>%
    summarise(count = sum(steps))

```
###Weekdays seem to be more active than weekends. 0-500 (5-minute) interval is dormant signaling sleep times. 
```{r}
 g <- ggplot(activity2_weekend,aes(interval, count))  
 g + geom_line(col="steelblue") + facet_grid(weekday ~ .)
  
 ```